More on the Slavnov products of spin chains, KP hierarchy tau and Baker-Akhiezer functions

### 2025-05-20 draft.tex
finish homogeneous limit

\date{\today}
\keywords{Integrability, Bethe states, Schur, Hall-Littlewood, Toda, KP}
\subjclass[2020]{82B20, 82B23}

@@

Given that each function \(\bm{\tau}^{(\ell)}(\bm{z}^{(\ell)}, \bm{v})\), for
\(\ell = 1, \dots, n+1\), is a distinct but related tau function, one
can define a vector
\begin{equation}
  \bm{T}(\bm{Z}, \bm{v}) = 
\begin{bmatrix}
\bm{\tau}^{(1)}(\bm{z}^{(1)}, \bm{v}) \\ 
\bm{\tau}^{(2)}(\bm{z}^{(2)}, \bm{v}) \\ 
\vdots \\ 
\bm{\tau}^{(n+1)}(\bm{z}^{(n+1)}, \bm{v})
\end{bmatrix}
\end{equation}
where \(\bm{Z} = (\bm{z}^{(1)}, \bm{z}^{(2)}, \dots,
\bm{z}^{(n+1)})\). This is a vector of tau functions of the KP
hierarchy. This observation suggests that there is, in fact, an
underlying multicomponent KP hierarchy in this context, and that the
linear system~(\ref{eq:linear-system}), which we can write as
\begin{equation}
\left( \bm{L} - \bm{\Lambda} \right) \bm{T} = \bm{0},
\end{equation}
defines a reduction of this multicomponent KP hierarchy. This
reduction ultimately describes constraints that relate the different
sets of parameters \(\bm{z}^{(\ell)}\).  We have not explored this
idea in this work, but we hope to return to this problem in future
publications.
\todo{16/05}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Tau functions expansion of the Slavnov product}
We can immediately observe that the poles of the tau
function~(\ref{eq:slav-tau-function}) are exactly the Bethe roots of
the integrable model.  We now want to analyse this pole structure of
\(\Delta(\bm{z})\tau(\bm{z}, \bm{v})\).  Let us consider that the
Bethe roots are not degenerate, that is, if \(j \neq k\), then \(v_j
\neq v_k\). In that case, the poles \(v_j\) are simple poles.

Let us now consider a particular coefficient \(\bm{z}\), say \(z_l\), and
make a Laplace expansion in the column \(l\), that is
\begin{equation}
\label{eq:minor-expansion}
\begin{split}
  \bm{\tau}(\bm{z}, \bm{v})
  & = \frac{1}{\Delta(\bm{z})} \det[\bm{\Omega}(\bm{z}) ] \\
  & = \frac{1}{\Delta(\bm{z})} \sum_{j=1}^n (-1)^{j + l} \frac{\mathcal{Y}_j(z_l)}{z_l - v_j  } 
  \det[\hat{\Omega}_{j, l}]\; ,
\end{split}
\end{equation}
where \(\hat{\Omega}_{j,l} \) is the \((j, l)\) minor of
\(\bm{\Omega}\).  Now, we can extract the residue of the tau function
for \(z_l\) at the point \(v_j\), that is
\begin{equation}
  \Res_{(z_l, v_j)}(\Delta(\bm{z})\tau(\bm{z}, \bm{v})) = (-1)^{j+l} \mathcal{Y}_j(v_j) \det[\hat{\Omega}_{j, l}]
\end{equation}

\begin{subequations}
Let us now decompose the Vandermonde determinant as
\begin{equation}
  \Delta(\bm{z}) = \prod_{j>k}(z_j - z_k) = \Delta(\bm{z}_l) \prod_{r<l}(z_l - z_r) \prod_{s>l}(z_s - z_l)\; ,
\end{equation}
where \(\bm{z}_l = \bm{z}\setminus \{z_l\} = (z_1, \dots, z_{l-1},
z_{l+1}, \dots, z_n) \), and we also define the function
\begin{equation}
  \Xi_l(z_l,\bm{z}_l) = \prod_{r<l}(z_l - z_r) \prod_{r>l}(z_r - z_l)\; , 
\end{equation}
then 
\begin{equation}
  \Delta(\bm{z}) = \Xi_l(z_l, \bm{z}_l) \Delta(\bm{z}_l) \; .
\end{equation}
\end{subequations}

@@

When we define the \(n\) sets of functions
\begin{equation}
  \begin{split}
    \hat{\bm{\Omega}}_{(j)}(z) & = \left(\Omega_1(z), \dots, \Omega_{j-1}(z), \Omega_{j+1}(z), \dots, \Omega_n(z) \right)\; ,
    \quad j =1, \dots, n\; ,
  \end{split}
\end{equation}
so that we can simplify the minor
expansion~(\ref{eq:minor-expansion-2}) even further; that is
\begin{equation}
\begin{split}
  \bm{\tau}(\bm{z}, \bm{v})
  & = \frac{1}{\Xi_l(z_l, \bm{z}_l)} \sum_{j=1}^n (-1)^{j + l} \frac{\mathcal{Y}_j(z_l)}{z_l - v_j} 
    \left(\frac{1}{\Delta(\bm{z}_l)} \det[\hat{\bm{\Omega}}_{(j) r}(\bm{z}_{(l) s} ]_{r, s=1}^n\right)\\ 
  & = \frac{1}{\Xi_l(z_l, \bm{z}_l)} \sum_{j=1}^n (-1)^{j + l} \frac{\mathcal{Y}_j(z_l)}{z_l - v_j} 
    \tilde{\bm{\tau}}_j(\bm{z}_l, \bm{v})\; ,
\end{split}
\end{equation}
where \(\hat{\bm{\Omega}}_{(j)r}\) is the \(r\)-th component of
\(\hat{\bm{\Omega}}_{(j)}\) and \(\hat{z}_{(l)s}\) is the \(s\)-th
component of \(\bm{z}_{(l)}\). Moreover, in the second line we have
defined the object
\begin{equation}
  \label{eq:basis-tau}
    \tilde{\bm{\tau}}_j(\bm{z}_l, \bm{v}) = 
\frac{1}{\Delta(\bm{z}_l)} \det[\hat{\bm{\Omega}}_{(j)}(\bm{z}_{(l)} ]\; .
\end{equation}
The curious aspect of this expression is that it easy to see that each term  
\(\tilde{\bm{\tau}}_j(\bm{z}_l, \bm{v})\) is a tau function itself, and that
it acts as a basis for the Slavnov's product. 
Additionally, these basis tau functions \(\tilde{\tau}_j(\bm{z}_l,
\bm{v})\) are, by construction, completely independent of \(z_l\).

Finally, we have
\begin{equation}
  \Res_{(z_l, v_j)}(\tau(\bm{z}, \bm{v}))
  = \frac{(-1)^{j + l} \mathcal{Y}_j(v_j)}{\Xi_l(v_j, \bm{z}_l)}
  \tilde{\bm{\tau}}_j(\bm{z}_l, \bm{v})
\end{equation}

In order to avoid cluttering, let us order the parameters \(\bm{z}\) and
consider these points close to the corresponding \(\bm{v}\), that is,
we consider just the residues \(z_j \to v_j\), then
\begin{subequations}
\begin{equation}
  \Res_{(z_j, v_j)}(\tau(\bm{z}, \bm{v}))
  = \frac{\mathcal{Y}_j(v_j)}{\Xi_j(v_j, \bm{z}_j)}
  \tilde{\bm{\tau}}_j(\bm{z}_j, \bm{v})
\end{equation}
or yet 
\begin{equation}
  \tilde{\bm{\tau}}_j(\bm{z}_j, \bm{v})
  = \frac{\Xi_j(v_j, \bm{z}_j)}{\mathcal{Y}_j(v_j)}
  \Res_{(z_j, v_j)}(\tau(\bm{z}, \bm{v}))
\end{equation}
\end{subequations}
Evidently, one can repeat the same ideas and take \(z_k \in \bm{z}_j\)
close to \(v_k \in \bm{v}_j\).

Therefore, we have the interesting result that the rresidues of the
tau function \(\tau(\bm{z}, \bm{v})\) for \(z_j\) approaching the
Bethe root \(v_j\) is proportional to a tau function.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Homogeneous limit}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

In the previous section we have discussed the limit where the
parameters \(\bm{z}\) approach the Bethe roots \(\bm{v}\).  We now
want to consider the limit where \(\bm{z}\) tend to a single variable,
say \(z_k\to z_1\), \(k=2,\dots, n\).

\subsection{Wronskian formula}

Let us first consider the case where \(z_2 \to z_1\). We consider a
series expansion about \(z_1\), and the second column of~(\ref{eq:slav-tau-function}) becomes
\begin{subequations}
\begin{equation}
  \bm{\tau}(\bm{z}, \bm{v}) = \frac{1}{\Delta(\bm{z})}
  \det
  \begin{pmatrix}
    \frac{\mathcal{Y}_1(z_1)}{z_1 - v_1}  & \frac{\mathcal{Y}_1(z_1)}{z_1 - v_1} + (z_2 - z_1) \frac{\mathcal{Y}_1^{(1)}(z_1)}{z_1 - v_1}
    + \mathcal{O}(\delta^2) &
    \frac{\mathcal{Y}_1(z_3)}{z_3 - v_1} & \dots & \frac{\mathcal{Y}_1(z_n)}{z_n - v_1}\\
    \frac{\mathcal{Y}_2(z_1)}{z_1 - v_2} & \frac{\mathcal{Y}_2(z_1)}{z_1 - v_2} + (z_2 - z_1) \frac{\mathcal{Y}_2^{(1)}(z_1)}{z_1 - v_2} 
    + \mathcal{O}(\delta^2) &
    \frac{\mathcal{Y}_2(z_3)}{z_3 - v_2} &  \dots & \frac{\mathcal{Y}_2(z_n)}{z_n - v_2}\\
    &  \vdots & & \\
    \frac{\mathcal{Y}_n(z_1)}{z_1 - v_n} & \frac{\mathcal{Y}_n(z_1)}{z_1 - v_n} + (z_2 - z_1) \frac{\mathcal{Y}_n^{(1)}(z_1)}{z_1 - v_n} 
    + \mathcal{O}(\delta^2) &
    \frac{\mathcal{Y}_n(z_3)}{z_3 - v_n} &  \dots & \frac{\mathcal{Y}_n(z_n)}{z_n - v_n}
  \end{pmatrix}
\end{equation}
where we have written \(z_2 - z_1 = \delta \to 0\), and we denote by
\(\mathcal{Y}^{(n)}(z)\) the \(n\)-th derivative of a
\(\mathcal{Y}(z)\) with respect its argument.

Now one can immediatelly see that the second column is the first
column plus terms multiplied by the factor \(\delta = z_2 -
z_1\). Using elementary operations on the column, we find
\begin{equation}
  \bm{\tau}_h(\bm{z}, \bm{v}) = \lim_{z_2 \to z_1}\frac{(z_2 - z_1)}{\Delta(\bm{z})}
  \det
  \begin{pmatrix}
    \frac{\mathcal{Y}_1(z_1)}{z_1 - v_1}  & \frac{\mathcal{Y}_1^{(1)}(z_1)}{z_1 - v_1} &
    \frac{\mathcal{Y}_1(z_3)}{z_3 - v_1} & \dots & \frac{\mathcal{Y}_1(z_n)}{z_n - v_1}\\
    \frac{\mathcal{Y}_2(z_1)}{z_1 - v_2} & \frac{\mathcal{Y}_2^{(1)}(z_1)}{z_1 - v_2} &
    \frac{\mathcal{Y}_2(z_3)}{z_3 - v_2} &  \dots & \frac{\mathcal{Y}_2(z_n)}{z_n - v_2}\\
    &  \vdots & & \\
    \frac{\mathcal{Y}_n(z_1)}{z_1 - v_n} & \frac{\mathcal{Y}_n^{(1)}(z_1)}{z_1 - v_n} &
    \frac{\mathcal{Y}_n(z_3)}{z_3 - v_n} &  \dots & \frac{\mathcal{Y}_n(z_n)}{z_n - v_n}
  \end{pmatrix}\; .
\end{equation}
We can now repeat the same ideas for \(z_3\to z_1=z_2\). We find that
the third column is the sum of the first and second columns, and
derivatives multiplied by the factor \(\delta^2\sim (z_3 - z_2)(z_3 -
z_1)\). All in all, we have
\begin{equation}
  \bm{\tau}_h(\bm{z}, \bm{v}) = \lim_{z_2, z_3 \to z_1}\frac{(z_2 - z_1)(z_3 - z_1)(z_3 - z_2)}{\Delta(\bm{z})}
  \det
  \begin{pmatrix}
    \frac{\mathcal{Y}_1(z_1)}{z_1 - v_1}  & \frac{\mathcal{Y}_1^{(1)}(z_1)}{z_1 - v_1} &
    \frac{\mathcal{Y}^{(2)}_1(z_1)}{z_1 - v_1} & \dots & \frac{\mathcal{Y}_1(z_n)}{z_n - v_1}\\
    \frac{\mathcal{Y}_2(z_1)}{z_1 - v_2} & \frac{\mathcal{Y}_2^{(1)}(z_1)}{z_1 - v_2} &
    \frac{\mathcal{Y}_2^{(2)}(z_1)}{z_1 - v_2} &  \dots & \frac{\mathcal{Y}_2(z_n)}{z_n - v_2}\\
    &  \vdots & & \\
    \frac{\mathcal{Y}_n(z_1)}{z_1 - v_n} & \frac{\mathcal{Y}_n^{(1)}(z_1)}{z_1 - v_n} &
    \frac{\mathcal{Y}_n^{(2)}(z_1)}{z_1 - v_n} &  \dots & \frac{\mathcal{Y}_n(z_n)}{z_n - v_n}
  \end{pmatrix}\; .
\end{equation}

Repeating the same idea for all columns, we get that the
multiplicative factor cancels out the Vandermonde determinant, and
that the homogeneous limit gives
\begin{equation}
  \bm{\tau}_h(z_1, \bm{v}) =
  \det
  \begin{pmatrix}
    \frac{\mathcal{Y}_1(z_1)}{z_1 - v_1}  & \frac{\mathcal{Y}_1^{(1)}(z_1)}{z_1 - v_1} &
    \frac{\mathcal{Y}^{(2)}_1(z_1)}{z_1 - v_1} & \dots & \frac{\mathcal{Y}_1^{(n-1)}(z_1)}{z_1 - v_1}\\
    \frac{\mathcal{Y}_2(z_1)}{z_1 - v_2} & \frac{\mathcal{Y}_2^{(1)}(z_1)}{z_1 - v_2} &
    \frac{\mathcal{Y}_2^{(2)}(z_1)}{z_1 - v_2} &  \dots & \frac{\mathcal{Y}_2^{(n-1)}(z_1)}{z_1 - v_2}\\
    &  \vdots & & \\
    \frac{\mathcal{Y}_n(z_1)}{z_1 - v_n} & \frac{\mathcal{Y}_n^{(1)}(z_1)}{z_1 - v_n} &
    \frac{\mathcal{Y}_n^{(2)}(z_1)}{z_1 - v_n} &  \dots & \frac{\mathcal{Y}^{(n-1)}_n(z_1)}{z_1 - v_n}
  \end{pmatrix}\; .
\end{equation}

Finally, let we write \(z_1 \equiv w\) and using elementary row operations we find
\begin{equation}
  \bm{\tau}_h(w, \bm{v}) =
  \frac{1}{\prod_{k=1}^n(w - v_k)}
  \mathcal{W}[\mathcal{Y}_1, \mathcal{Y}_2, \dots, \mathcal{Y}_n](w)
  \; .
\end{equation}
\end{subequations}
The interesting aspect of this expression is that the poles of this
function are exactly the Bethe roots, and they are easily built from
the Algebraic Bethe ansatz. Additionally, the functions
\(\{\mathcal{Y}_j(w)\}_{j=1}^n\) are linearly independent if and only
if the Bethe roots are non-degenerate.

>>>>

\date{\today}
\keywords{Integrability, spin chains, bethe, slavnov, kp}
\subjclass[2020]{82B20, 82B23}

@@

Given that each function \(\bm{\tau}^{(\ell)}(\bm{z}^{(\ell)}, \bm{v})\), for
\(\ell = 1, \dots, n+1\), is a distinct but related tau function, one
can define a vector
\begin{equation}
  \bm{T}(\bm{Z}, \bm{v}) = 
\begin{bmatrix}
\bm{\tau}^{(1)}(\bm{z}^{(1)}, \bm{v}) \\ 
\bm{\tau}^{(2)}(\bm{z}^{(2)}, \bm{v}) \\ 
\vdots \\ 
\bm{\tau}^{(n+1)}(\bm{z}^{(n+1)}, \bm{v})
\end{bmatrix}
\end{equation}
where \(\bm{Z} = (\bm{z}^{(1)}, \bm{z}^{(2)}, \dots,
\bm{z}^{(n+1)})\). This is a vector of tau functions of the KP
hierarchy. This observation suggests that there is, in fact, an
underlying multicomponent KP hierarchy in this context, and that the
linear system~(\ref{eq:linear-system}), which we can write as
\begin{equation}
\left( \bm{L} - \bm{\Lambda} \right) \bm{T} = \bm{0},
\end{equation}
defines a reduction of this multicomponent KP hierarchy. This
reduction ultimately describes constraints that relate the different
sets of parameters \(\bm{z}^{(\ell)}\).  We have not explored this
idea in this work, but we hope to return to this problem in future
publications.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Tau functions expansion of the Slavnov product}

Let us now fix a component \(\ell\) and omit this index in
equation~(\ref{eq:slav-tau-function}). We also assume that the Bethe
roots are non-degenerate; that is, for \(j \neq k\), we necessarily
have \(v_j \neq v_k\). In other words, each Bether root \(v_j\)
corresponds to a simple poles of the Slavnov product.

Fix a particular coefficient of \(\bm{z}\), say \(z_l\), and consider a Laplace
expansion of~(\ref{eq:slav-tau-function}) along the \(l\)-th column. That is
\begin{equation}
\label{eq:minor-expansion}
\begin{split}
  \bm{\tau}(\bm{z}, \bm{v})
  & = \frac{1}{\Delta(\bm{z})} \det[\bm{\Omega}(\bm{z}) ] \\
  & = \frac{1}{\Delta(\bm{z})} \sum_{j=1}^n (-1)^{j + l} \frac{\mathcal{Y}_j(z_l)}{z_l - v_j  } 
  \det[\hat{\Omega}_{j, l}]\; ,
\end{split}
\end{equation}
where \(\det[\hat{\Omega}_{j,l}]\) denotes the \((j, l)\)-minor of
\(\bm{\Omega}\). We can now extract the residue of the tau function
with respect to \(z_l\) at the point \(v_j\); that is
\begin{equation}
  \Res_{z_l=v_j}(\Delta(\bm{z})\tau(\bm{z}, \bm{v})) = (-1)^{j+l} \mathcal{Y}_j(v_j) \det[\hat{\Omega}_{j, l}]\;. 
\end{equation}

\begin{subequations}
Additionally, decompose the Vandermonde determinant as
\begin{equation}
  \Delta(\bm{z}) = \prod_{j>k}(z_j - z_k) = \Delta(\bm{z}_l) \prod_{r<l}(z_l - z_r) \prod_{s>l}(z_s - z_l)\; ,
\end{equation}
where \(\bm{z}_l = \bm{z}\setminus \{z_l\} = (z_1, \dots, z_{l-1},
z_{l+1}, \dots, z_n) \), and we also define the function
\begin{equation}
  \Xi_l(z_l,\bm{z}_l) = \prod_{r<l}(z_l - z_r) \prod_{r>l}(z_r - z_l)\; , 
\end{equation}
then 
\begin{equation}
  \Delta(\bm{z}) = \Xi_l(z_l, \bm{z}_l) \Delta(\bm{z}_l) \; .
\end{equation}
\end{subequations}

@@

By defining the \(n\) sets of functions
\begin{equation}
  \begin{split}
    \hat{\bm{\Omega}}_{(j)}(z) & = \left(\Omega_1(z), \dots, \Omega_{j-1}(z), \Omega_{j+1}(z), \dots, \Omega_n(z) \right)\; ,
    \quad j =1, \dots, n\; , 
  \end{split}
\end{equation}
we can further simplify the minor
expansion~(\ref{eq:minor-expansion-2}); that is
\begin{equation}
  \label{eq:slavnov-tauexp}
\begin{split}
  \bm{\tau}(\bm{z}, \bm{v})
  & = \frac{1}{\Xi_l(z_l, \bm{z}_l)} \sum_{j=1}^n (-1)^{j + l} \frac{\mathcal{Y}_j(z_l)}{z_l - v_j} 
    \left(\frac{1}{\Delta(\bm{z}_l)} \det[\hat{\bm{\Omega}}_{(j) r}(\bm{z}_{(l) s} ]_{r, s=1}^n\right)\\ 
  & = \frac{1}{\Xi_l(z_l, \bm{z}_l)} \sum_{j=1}^n (-1)^{j + l} \frac{\mathcal{Y}_j(z_l)}{z_l - v_j} 
    \tilde{\bm{\tau}}_j(\bm{z}_l, \bm{v})\; ,
\end{split}
\end{equation}
where \(\hat{\bm{\Omega}}_{(j)r}\) is the \(r\)-th component of
\(\hat{\bm{\Omega}}_{(j)}\) and \(\hat{z}_{(l)s}\) is the \(s\)-th
component of \(\bm{z}_{(l)}\). Moreover, in the second line we have
defined the object
\begin{equation}
  \label{eq:basis-tau}
    \tilde{\bm{\tau}}_j(\bm{z}_l, \bm{v}) = 
\frac{1}{\Delta(\bm{z}_l)} \det[\hat{\bm{\Omega}}_{(j)}(\bm{z}_{(l)})]\; .
\end{equation}
Finally, we have
\begin{equation}
  \Res_{z_l = v_j}(\tau(\bm{z}, \bm{v}))
  = \frac{(-1)^{j + l} \mathcal{Y}_j(v_j)}{\Xi_l(v_j, \bm{z}_l)}
  \tilde{\bm{\tau}}_j(\bm{z}_l, \bm{v})\; .
\end{equation}

We can also organize the parameters \(\bm{z}\) and consider these
points close to the corresponding Bethe roots \(\bm{v}\). Therefore 
\begin{subequations}
\begin{equation}
  \Res_{z_j = v_j}(\tau(\bm{z}, \bm{v}))
  = \frac{\mathcal{Y}_j(v_j)}{\Xi_j(v_j, \bm{z}_j)}
  \tilde{\bm{\tau}}_j(\bm{z}_j, \bm{v})\; ,
\end{equation}
or yet 
\begin{equation}
  \tilde{\bm{\tau}}_j(\bm{z}_j, \bm{v})
  = \frac{\Xi_j(v_j, \bm{z}_j)}{\mathcal{Y}_j(v_j)}
  \Res_{z_j = v_j}(\tau(\bm{z}, \bm{v}))\; .
\end{equation}
\end{subequations}

Expression~(\ref{eq:slavnov-tauexp}) is one of the main results in
this work.  It is easy to see that each term
\(\tilde{\bm{\tau}}_j(\bm{z}_l, \bm{v})\) is a tau function itself,
and serves as a basis for the Slavnov's product. Additionally, these
basis tau functions \(\tilde{\bm{\tau}}_j(\bm{z}_l, \bm{v})\) are, by
construction, completely independent of \(z_l\). This result
essentially shows that, given a \(\mathcal{Y}\)-function (which is
related to the eigenvalue of the transfer matrix), one can construct a
basis of tau functions that span the Slavnov products in the
corresponding integrable system. Moreover, this expansion ensures that
the resulting Slavnov product is also a tau function of the KP
hierarchy.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Homogeneous limit: Wronskian formula}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

In the previous section, we discussed the limit in which the
parameters \(\bm{z}\) approach the Bethe roots \(\bm{v}\).  We now
consider a different limit, where all variables in the set \(\bm{z}\)
tend to a single variable -- that is, \(z_k\to z_1\) for \(k=2,\dots,
n\).

Let us first consider the case \(z_2 \to z_1\). We perform a series
expansion around \(z_1\), in which the second column
of~(\ref{eq:slav-tau-function}) becomes
\begin{subequations}
\begin{equation}
  \bm{\tau}(\bm{z}, \bm{v}) = \frac{1}{\Delta(\bm{z})}
  \det
  \begin{pmatrix}
    \frac{\mathcal{Y}_1(z_1)}{z_1 - v_1}  & \frac{\mathcal{Y}_1(z_1)}{z_1 - v_1} + (z_2 - z_1) \frac{\mathcal{Y}_1^{(1)}(z_1)}{z_1 - v_1}
    + \mathcal{O}(\delta^2) &
    \frac{\mathcal{Y}_1(z_3)}{z_3 - v_1} & \dots & \frac{\mathcal{Y}_1(z_n)}{z_n - v_1}\\
    \frac{\mathcal{Y}_2(z_1)}{z_1 - v_2} & \frac{\mathcal{Y}_2(z_1)}{z_1 - v_2} + (z_2 - z_1) \frac{\mathcal{Y}_2^{(1)}(z_1)}{z_1 - v_2} 
    + \mathcal{O}(\delta^2) &
    \frac{\mathcal{Y}_2(z_3)}{z_3 - v_2} &  \dots & \frac{\mathcal{Y}_2(z_n)}{z_n - v_2}\\
    &  \vdots & & \\
    \frac{\mathcal{Y}_n(z_1)}{z_1 - v_n} & \frac{\mathcal{Y}_n(z_1)}{z_1 - v_n} + (z_2 - z_1) \frac{\mathcal{Y}_n^{(1)}(z_1)}{z_1 - v_n} 
    + \mathcal{O}(\delta^2) &
    \frac{\mathcal{Y}_n(z_3)}{z_3 - v_n} &  \dots & \frac{\mathcal{Y}_n(z_n)}{z_n - v_n}
  \end{pmatrix}\; ,
\end{equation}
where we have written \(z_2 - z_1 = \delta \to 0\). Moreover, let us
denote by \(\mathcal{Y}^{(n)}(z)\) the \(n\)-th derivative of
\(\mathcal{Y}(z)\) with respect its argument.

One can immediately see that the second column is equal to the first
column plus terms proportional to the factor \(\delta = z_2 -
z_1\)​. Using elementary column operations, we find
\begin{equation}
  \bm{\tau}_h(\bm{z}, \bm{v}) = \lim_{z_2 \to z_1}\frac{(z_2 - z_1)}{\Delta(\bm{z})}
  \det
  \begin{pmatrix}
    \frac{\mathcal{Y}_1(z_1)}{z_1 - v_1}  & \frac{\mathcal{Y}_1^{(1)}(z_1)}{z_1 - v_1} &
    \frac{\mathcal{Y}_1(z_3)}{z_3 - v_1} & \dots & \frac{\mathcal{Y}_1(z_n)}{z_n - v_1}\\
    \frac{\mathcal{Y}_2(z_1)}{z_1 - v_2} & \frac{\mathcal{Y}_2^{(1)}(z_1)}{z_1 - v_2} &
    \frac{\mathcal{Y}_2(z_3)}{z_3 - v_2} &  \dots & \frac{\mathcal{Y}_2(z_n)}{z_n - v_2}\\
    &  \vdots & & \\
    \frac{\mathcal{Y}_n(z_1)}{z_1 - v_n} & \frac{\mathcal{Y}_n^{(1)}(z_1)}{z_1 - v_n} &
    \frac{\mathcal{Y}_n(z_3)}{z_3 - v_n} &  \dots & \frac{\mathcal{Y}_n(z_n)}{z_n - v_n}
  \end{pmatrix}\; .
\end{equation}
We can now repeat the same reasoning for \(z_3\to z_1=z_2\). We find
that the third column becomes a linear combination of the first and
second columns, along with terms involving derivatives multiplied by
the factor \(\delta^2 = (z_3 - z_2)(z_3 - z_1)\). All in all, we have
\begin{equation}
  \bm{\tau}_h(\bm{z}, \bm{v}) = \lim_{z_2, z_3 \to z_1}\frac{(z_2 - z_1)(z_3 - z_1)(z_3 - z_2)}{\Delta(\bm{z})}
  \det
  \begin{pmatrix}
    \frac{\mathcal{Y}_1(z_1)}{z_1 - v_1}  & \frac{\mathcal{Y}_1^{(1)}(z_1)}{z_1 - v_1} &
    \frac{\mathcal{Y}^{(2)}_1(z_1)}{z_1 - v_1} & \dots & \frac{\mathcal{Y}_1(z_n)}{z_n - v_1}\\
    \frac{\mathcal{Y}_2(z_1)}{z_1 - v_2} & \frac{\mathcal{Y}_2^{(1)}(z_1)}{z_1 - v_2} &
    \frac{\mathcal{Y}_2^{(2)}(z_1)}{z_1 - v_2} &  \dots & \frac{\mathcal{Y}_2(z_n)}{z_n - v_2}\\
    &  \vdots & & \\
    \frac{\mathcal{Y}_n(z_1)}{z_1 - v_n} & \frac{\mathcal{Y}_n^{(1)}(z_1)}{z_1 - v_n} &
    \frac{\mathcal{Y}_n^{(2)}(z_1)}{z_1 - v_n} &  \dots & \frac{\mathcal{Y}_n(z_n)}{z_n - v_n}
  \end{pmatrix}\; .
\end{equation}

By applying this procedure iteratively to each column, we find that
the multiplicative factors cancel the Vandermonde determinant, and the
homogeneous limit becomes
\begin{equation}
  \bm{\tau}_h(z_1, \bm{v}) =
  \det
  \begin{pmatrix}
    \frac{\mathcal{Y}_1(z_1)}{z_1 - v_1}  & \frac{\mathcal{Y}_1^{(1)}(z_1)}{z_1 - v_1} &
    \frac{\mathcal{Y}^{(2)}_1(z_1)}{z_1 - v_1} & \dots & \frac{\mathcal{Y}_1^{(n-1)}(z_1)}{z_1 - v_1}\\
    \frac{\mathcal{Y}_2(z_1)}{z_1 - v_2} & \frac{\mathcal{Y}_2^{(1)}(z_1)}{z_1 - v_2} &
    \frac{\mathcal{Y}_2^{(2)}(z_1)}{z_1 - v_2} &  \dots & \frac{\mathcal{Y}_2^{(n-1)}(z_1)}{z_1 - v_2}\\
    &  \vdots & & \\
    \frac{\mathcal{Y}_n(z_1)}{z_1 - v_n} & \frac{\mathcal{Y}_n^{(1)}(z_1)}{z_1 - v_n} &
    \frac{\mathcal{Y}_n^{(2)}(z_1)}{z_1 - v_n} &  \dots & \frac{\mathcal{Y}^{(n-1)}_n(z_1)}{z_1 - v_n}
  \end{pmatrix}\; .
\end{equation}
Finally, we write \(z_1 \equiv w\) and using elementary row operations
we find
\begin{equation}
  \bm{\tau}_h(w, \bm{v}) =
  \frac{1}{\prod_{k=1}^n(w - v_k)}
  \mathcal{W}[\mathcal{Y}_1, \mathcal{Y}_2, \dots, \mathcal{Y}_n](w)
  \; .
\end{equation}
\end{subequations}
An interesting aspect of this expression is that its poles coincide
precisely with the Bethe roots \(\bm{v}\). Moreover, the homogeneous
limit can be readily constructed from the Algebraic Bethe Ansatz,
since it depends only on the transfer matrix
eigenvalues. Additionally, the functions
\(\{\mathcal{Y}_j(w)\}_{j=1}^n\) are linearly independent if and only
if the Bethe roots are non-degenerate.
